---
title: "Beginners' Handbook of the `EMVportfolio` R package"
author: Yuling Max Chen
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
  pdf_document:
    number_sections: true
    toc: TRUE
    toc_depth: 3
    citation_package: natbib
bibliography: Reference.bib
biblio-style: apalike
reference-section-title: Bibliography
vignette: >
  %\VignetteIndexEntry{Introduction `trimFZC` R package}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72 
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, eval=TRUE)
```

The `EMVportfolio` package is designed to solve for the Exploratory Mean-Variance (EMV) portfolio optimization problem, as described in [@wang2020continuous]. In this document, we illustrate the main features of this R package with examples. 


# Background Information
Exploratory Mean-Variance (EMV) portfolio optimization problem [@wang2020continuous] is an investment problem, that aims to maximize the expected portfolio value (wealth) while bearing the minimal risk (variance). According to [@wang2020continuous], the EMV problem is formulated under a Reinforcement Learning (RL) framework, and the optimal exploratory policy follows a Gaussian distribution, from which we can sample the investment decisions (i.e., the dollar value invested in the stock). By parameterizing the mean $\mu$ and standard-deviation $\sigma$ of the Gaussian policy distribution, [@wang2020continuous] constructs a RL training loss, named as the Temporal Difference (TD) error. And the Gaussian policy distribution can be optimized by iteratively updating the parameters and minimizing the TD error. 

For comparison purpose, [@wang2020continuous] also introduced the classical (non-exploratory) Mean-Variance (MV) portfolio optimization problem, which can be solved via Maximum Likelihood Estimation (MLE). 

For more details of the EMV problem, please refer to [@wang2020continuous]. 


# Installation

One can download the latest version of this package from github: 
```{r eval=FALSE, message=FALSE}
require(devtools)
devtools::install_github("MaxGniluynehc/EMVportfolio",
               ref="main",
               auth_token = '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',
               build_vignettes = TRUE)
library(EMVportfolio)
library(zoo)
```


# Overview 

The `EMVportfolio` package has the following architecture: 

```{r echo=FALSE, eval=FALSE, include=FALSE}
PATH = "/Users/maxchen/Documents/Study/STA/STAT900/Writings/Final Project/showdir"
fs::dir_tree(path=PATH, type="any")
```

```{r}
# └── R
#     ├── act_EMV_policy.R
#     ├── act_MLE_policy.R
#     ├── getStockPath.R
#     ├── getTerminalWealth.R
#     ├── show_terminalwealth.R
#     └── trainRLalg.R
```

* The `act_EMV_policy.R` and `act_MLE_policy.R` contains functions that computes the investment decision/action (i.e., dollar value invested in the stock), following either the classical policy with MLE, or the exploratory EMV policy under the RL framework. 
* The `getStockPath.R` simulates a path of stock price process. 
* The `getTerminalWealth.R` trades based on a given policy and computes the terminal wealth, i.e., the portfolio value at the end of the investment period. 
* The `show_terminalwealth.R` visualizes the simulations of terminal wealth with plots and tables.  
* The `trainRLalg.R` optimizes the exploratory policy by training a parameterized RL model and minimizing the TD error. It also contains plotting functions to illustrate the convergence of the training process. 



# Examples

## Simulating Stock Price Paths

One can easily simulate stock price paths by: 
```{r}
set.seed(1234)
N_sim = 10
for (i in 1:N_sim){
  St= getStockPath(mu=-0.3, sigma=0.2, mwindow_size = 0)
  if (i==1){
    par(mai=c(1.02,1,0.4,0.3))
    plot(St, col="gray", type="l", ylim = c(0.3,1.5),
         ylab="Stock Price ($)", xlab = "Time (days)",
         cex.lab=1.5)
  }
  else{
    lines(St, col="gray")
  }
}
```



## Training the RL Model

Training the RL model is simply one line of code, with all arguments taking their default values: 
```{r eval=FALSE}
# For time and presentation purpose, this code chunck is stopped. One can manually set `eval = TRUE` to let it run. 
trainRLalg_output = trainRLalg()
```

The default arguments are chosen following the simulation study in [@wang2020continuous], and it guarantees a converged training path. One can also try to input different values to this function, and see how the convergence of the learning process varies. 

For example, one can change the learning rates (`alpha, lr_theta, lr_phi`) to larger values and let each training epoch update a larger step along the gradient. One can also encourage larger updating steps by setting `clip_gradient` to `NULL`, i.e., remove the restriction on the gradient computation. Besides, one can also manually choose the initial points for the parameters (`theta_0`, `theta_1`, `theta_2`, `theta_3`, `phi1`, `phi2`, `w`). 

To change the market scenario, one can consider changing (`mu`, `sigma`, `S0`, `rf`) to different values. Or, to change the investment period and the continuous-time discretization, one can input different values to (`invest_hrzn`, `dt`). 

The training out, `trainRLalg_output`, is a named list containing the parameters and the training loss (TD error) over the training epoch. For details of what information does `trainRLalg_output` contain, please see the R helper documentation by `help(trainRLalg)`. 


### Showing Training Result
Suppose after training the RL model, we get `trainRLalg_output`. Then we can visualize the training procedure by plotting the paramsters and the training loss (TD error). 

```{r echo=FALSE}
PATH = paste0(getwd(), "/train_result_demo/")
params = read.csv(paste0(PATH, "params.csv"))
TD_errors = read.csv(paste0(PATH, "TD_errors.csv"))
trainRLalg_output = list("params"=params[,2:ncol(params)], "TD_errors"=TD_errors)
```


```{r}
par(mai=c(1.02,1,0.4,0.3))
plot_trainResult(trainRLalg_output, type = "params", cex.lab=1.5)

par(mai=c(1.02,1,0.4,0.3))
plot_trainResult(trainRLalg_output, type = "thetas", cex.lab=1.5)

par(mai=c(1.02,1,0.4,0.3))
plot_trainResult(trainRLalg_output, type = "phis", cex.lab=1.5)

par(mai=c(1.02,1,0.4,0.3))
plot_trainResult(trainRLalg_output, type = "TD_errors", cex.lab=1.5)

```


Moreover, the functions adapts most (if not all) the embedded arguments of the basic `plot()` function, e.g., to change the limit of the y-axis and change the lable on the x-axis, one can try: 

```{r eval=FALSE}
plot_trainResult(trainRLalg_output, type = "params", ylim=c(-5,5), 
                 xlab = "Some other x-axis label")
```


Upon convergence, it is recommended to save the trained (and optimized) parameters to a local directory, so that we can recycle them and conduct simulations. 


## Conduct a Simulation 

According to [@wang2020continuous], there are 2 major investment polices: the classical Mean-Variance (MV) policy and the Exploratory Mean-Variance (EMV) policy. The classical MV policy is a deterministic function, and the EMV policy is a Gaussian distribution. Moreover, the classical MV policy is exactly the mean of the EMV policy. For the 2 policies, one can either directly use MLE to estimate the parameters in the policy, or use the RL algorithm to train the reparameterized policy. Moreover, since the MLE approach is known to suffer the **mean-blur problem**, we place a constraint on the investment decisions, in order to produce more stable outputs. 

Therefore, we consider 4 different trading strategies: 
1. Unconstrained MLE approach on the classical MV policy; 
2. Constrained MLE approach on the classical MV policy, with the maximum investment a 
trader can make on a single day is restricted to [-5, +5];
3. Constrained MLE approach on the EMV policy, with the maximum investment a 
trader can make on a single day is restricted to [-5, +5];
4. Unconstrained EMV policy optimized by the RL algorithm. 

For each trading strategy, we consider to simulate 5000 terminal wealths, each with `N_runs=50` runs and each run has `N_sim=100` simulations. The code to conduct this simulation is as follows:  

```{r eval=FALSE}
# For time purpose, this code chunck is not executed, but one can set `eval=TRUE` to run it. 
N_runs = 50
N_sim = 100
set.seed(1111)

phi1 = params[nrow(params),"phi1"]
phi2 = params[nrow(params),"phi2"]
w = params[nrow(params), "w"]

terminal_wealths = list("MLE_classic_u"=c(), "MLE_classic"=c(),
                        "MLE_exploratory"=c(), "EMV"=c())
for(run in 1:N_runs){
  MLEcu = c()
  MLEc = c()
  MLEe = c()
  EMV = c()
  for(n in 1:N_sim){
    print(paste0("run: ", run, " in ", N_runs, "; sim: ", n, " in ", N_sim))
    St = getStockPath(mu=-0.3, sigma=0.2, mwindow_size = 100)
    Out_MLEcu = getTerminalWealth(St, "MLE classic", constrained_invest = F)
    Out_MLEc = getTerminalWealth(St, "MLE classic", constrained_invest = T, max_invest = 5)
    Out_MLEe = getTerminalWealth(St, "MLE exploratory", constrained_invest = T, max_invest = 5)
    MLEcu = c(MLEcu, Out_MLEcu$terminal_wealth)
    MLEc = c(MLEc, Out_MLEc$terminal_wealth)
    MLEe = c(MLEe, Out_MLEe$terminal_wealth)

    St_emv = St[101: length(St)]
    Out_EMV = getTerminalWealth(St_emv, "EMV", phi1=phi1, phi2=phi2, w=w)
    EMV = c(EMV, Out_EMV$terminal_wealth)
  }
  terminal_wealths$MLE_classic_u = cbind(terminal_wealths$MLE_classic_u, MLEcu)
  terminal_wealths$MLE_classic = cbind(terminal_wealths$MLE_classic, MLEc)
  terminal_wealths$MLE_exploratory = cbind(terminal_wealths$MLE_exploratory, MLEe)
  terminal_wealths$EMV = cbind(terminal_wealths$EMV, EMV)
}

```


## Visualizing Simualtion Results

Once we've done the simulation, we will obtain a named list `terminal_wealths` containing the terminal wealth following the 4 trading strategies.  


```{r echo=FALSE}
terminal_wealths = list("MLE_classic_u"=read.csv2(paste0(PATH, "MLE_classic_u.csv"))[, -1], 
                        "MLE_classic"=read.csv2(paste0(PATH, "MLE_classic.csv"))[, -1],
                        "MLE_exploratory"=read.csv2(paste0(PATH,"MLE_exploratory.csv"))[, -1],
                        "EMV"=read.csv2(paste0(PATH, "EMV.csv"))[, -1])

```

```{r}
names(terminal_wealths)
```
### Tabularizing Results
Then, we can present a summary table showing the quality of the simulations, using the `table_terminalwealth()` funtion: 
```{r}
MLE = table_terminalwealth(terminal_wealths$MLE_classic_u, "summary")
MLE_constrained = table_terminalwealth(terminal_wealths$MLE_classic, "summary")
MLE_exploratory = table_terminalwealth(terminal_wealths$MLE_exploratory, "summary")
EMV = table_terminalwealth(terminal_wealths$EMV, "summary")
knitr::kable(data.frame(rbind(MLE, MLE_constrained, MLE_exploratory, EMV)))

```


We can also show the quantiles of the simulations, for comparison purpose: 


```{r}
MLE_q = table_terminalwealth(unlist(terminal_wealths$MLE_classic_u), "quantile")
MLE_constrained_q = table_terminalwealth(unlist(terminal_wealths$MLE_classic), "quantile")
MLE_exploratory_q = table_terminalwealth(unlist(terminal_wealths$MLE_exploratory), "quantile")
EMV_q = table_terminalwealth(unlist(terminal_wealths$EMV), "quantile")
knitr::kable(data.frame(round(rbind(MLE_q, MLE_constrained_q, MLE_exploratory_q, EMV_q), digits = 3)))
```

For details of the `table_terminalwealth()` function, please refer to the R documentation by `help(table_terminalwealth)`. 


### Plotting Results
Apart from the tables, we can draw 4 types of plots: the Boxplot, Density plot, Empirical CDF (ECDF) plot, and the QQplot, using the `plot_terminalwealth()` function. This can be done by simply setting the `plot_type` to one of (`boxplot`, `density`, `ecdf`, `qqplot`). 

```{r}
tws = cbind(unlist(terminal_wealths$MLE_classic_u),
           unlist(terminal_wealths$MLE_classic),
           unlist(terminal_wealths$MLE_exploratory),
           unlist(terminal_wealths$EMV))
colnames(tws) = c("MLE_classic_u", "MLE_constrained", "MLE_exploratory", "RL")

par(mai=c(1.02,1,0.4,0.3))
plot_terminalwealth(tws, plot_type = "boxplot",
                    ylim=c(-3,5), 
                    ylab="Terminal Wealth ($)", cex.lab=1.5)


par(mai=c(1.02,1,0.4,0.3))
plot_terminalwealth(tws, plot_type = "density",
                    xlim=c(min(terminal_wealths$MLE_classic ),
                           max(terminal_wealths$MLE_exploratory)),
                    ylim=c(0,6), cex.lab=1.5)


par(mai=c(1.02,1,0.4,0.3))
plot_terminalwealth(tws, plot_type = "ecdf",
                    xlim=c(-3,5), ylim=c(0,1), cex.lab=1.5)

```


For QQplot, one thing to note is that we can only input **a pair** of simulations, i.e., the column number of the input matrix has to be 2. Because QQplot only compares the quantiles between 2 sequences. So, we can write a *double for-loop* to compare all the pairs of simulations: 
```{r fig.height=10}
par(mai=c(1.02,1,0.4,0.3), mfrow=c(3,2))
for (i in 1:(ncol(tws)-1)){
  for (j in (i+1):ncol(tws)){
    if(i == 1){
      plot_terminalwealth(tws[,c(i,j)], plot_type = "qqplot",
                          xlim=c(-1000,1500), cex.lab=2)
    }
    else{
      plot_terminalwealth(tws[,c(i,j)], plot_type = "qqplot", cex.lab=2)
    }
  }
}
```

For details of the `plot_terminalwealth()` function, please refer to the R documentation by `help(plot_terminalwealth)`. 

Again, the `plot_terminalwealth()` function adapts the embedded arguments for the basic `plot()` function, e.g., one can change the range of the y-axis by giving a different value for `ylim`. 



# Package Testings

## R CMD Checks
As a general checking of the correct design and structure of this package, it is recommended to do conduct a R CMD Check, which can be simply done via the `check()` function in `devtools` package, i.e.,  `devtools::check()`. 

Currently our package passes all the R CMD Checks. 


## Testing for Invalid Inputs
All the packages 









